---
title: "Exercise: Data Management 1"
output: html_document
---

```{r setup, include=FALSE}
course_dir <- file.path("C:", "Users", "rdefa", "rdefaccio_outlook.com", 
                        "OneDrive", "Desktop", "1a - Active Projects", 
                        "CDC Phase 2 R Dec. 2022", "R_Programming_CDC_Bootcamp")
knitr::opts_knit$set(warning = FALSE, message = FALSE, root.dir = course_dir)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, root.dir = course_dir)
```

## Data Management Exercise (45-60 minutes)

In the following exercise, you'll bring in some data and begin the process of cleaning it. The data files used for this and following exercises are located in the data/samples folder, and are in a SAS format.

-   `data/samples/lb.sas7bdat`
-   `data/samples/dm.sas7bdat`

## Bring in the `lb` file

Import the `lb` file

-   What format is the file? What reader function can you use to bring it in?
-   Look at the head of the file.
-   Look at the structure of the file using `str()` or `tibble::glimpse()`

```{r}

```

## Examine the meta-information

Because this is a SAS file, `haven` imported some additional information as "attributes" that include longer titles and SAS formats.

You can see the information in the `str()` results or in the `glimpse()` results

To pull out the attributes for a specific column, try using the `attr()` function. The syntax should be like follows:

> `attr(dataframe, "attribute_name")`

Try to pull up the attribute "label" for the first column in `lb`

```{r}

```

See if there is a "format.sas" label for that same column.

```{r}

```

Now try adapting the following `lapply()` ("list apply") command to see attributes for all the columns.

`lapply(dataframe, attr, "label")`

-   `dataframe` should be the name of your data

-   `attr` is the `attr()` function which can set or get an attribute

-   `"label"` is the label of the specific attributes.

First try with the "label" attribute.

```{r}

```

Now try with the "format.sas" attribute.

```{r}

```

## Change the SUBJID data type

-   What is the current `SUBJID` data type? Change it to character.

```{r}

```

## Change EPOCH to factor

-   create a `table()` of the variable EPOCH

-   how many `unique()` values does EPOCH contain

-   convert EPOCH to a factor

    -   What will you do with any blank ("") labels?

    -   Check the factored variable with a `table()` command after conversion. Did you accomplish what you intended?

```{r}


```

## Change other factors

-   What other variables could be consider factors?

-   If it make sense, change them into factors.

    -   You can always create a factored copy of a variable and keep the original, to give you flexibility.

```{r}

```

## Recode LBBLFL

-   The LBBLFL has values "Y" or "". Recode it into a factor that has 0 and 1 values instead, ("Y" =1, "" = 0.)

-   Use both the `levels=` and `labels=` options of the `factor()` function

-   Check with a `table()` afterwards

```{r}

```

## Missing data

-   Check the LBSTNRLO and LBSTNRHI for missing data.

-   How many missing values are there in each variable?

-   What is the percent missing for each variable?

```{r}

```

## Convert LBDTC to a date

-   Convert LBDTC to a date time.

    -   First inspect the data. What format is it in: year first or year last? Where are the month and day components? What is the separator?

    -   Apply a lubridate command to the variable.

    -   Check your conversion. Did you do what you wanted?

```{r}

```

## Using `ifelse` to recode data

The lab results sometimes have lower and upper reference ranges. Let's create a flag that indicates if the lab results is **within** this range.

-   Use the `ifelse()` function to create an indicator variable called LB_IN_RANGE

-   The logic of the variable should be as such:

    -   TRUE if the value of the lab variable is above the lower reference value and below the upper reference value.

    -   FALSE otherwise.

    -   use the code pattern VAR = ifelse( comparison_statement, TRUE, FALSE)

        -   You will have to change the VAR name and figure out the comparison_statement logic

    -   Check your resulting variable with a table() command. Does it make sense?

```{r}

```

## Save your cleaned results

Are there any other changes you would make to the data?

In this exercise, you may have copied variables as you transformed them. Do you want to keep the original raw data or eliminate it from the data set.

-   If you want to drop a variable or column from the data set, you can set it to NULL.

    -   `dta$col <- NULL`

-   Or, you may choose to keep the original "raw" variable.

    -   If you keep both copies, someone may get confused about which is the "correct" variable, or inadvertently use the wrong one.

    -   On the other hand, the original copy provides a "backup" variable.

    -   Talk with your colleagues about best practices, and about naming conventions.

-   Once you have arrived at your finalized "clean" data set, save it.

    -   You can use the `readr::write_csv()` command to write it to disk.

    -   Or you could use other write commands to write to different formats. Look at `?readr` for more details. For even more import/export options, look at the `rio` (<https://github.com/leeper/rio>) package.

```{r}
#this will write the output into the data/samples folder. 

```

Also try saving this to a *.rds* file.

```{r}

```
